Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees

### 2023-02-20 script/manuscript_MEE.Rmd
working on expert reclass

@@

@@

# all samples with duplicate genera in them # be careful these are not easy mappings
duplicate_reps_in_sample <- blst_expert %>%

  # identify the duplicate samples
  drop_na(plant.species) %>% 
  group_by(sample.id, TAXON_NEW) %>% 
  ungroup(TAXON_NEW)  %>% 
  distinct(plant.species, .keep_all = T) %>% 
  group_by(sample.id, TAXON_NEW) %>% 
  filter(n() >= 2) %>%

  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Soloman'
  ungroup() %>% 
  group_by(RecordID) %>% 
  mutate(Prcnt_seqs_Reclass = Prcnt_seqs/n())

blst_reclass_expert <- blst_expert %>% 
  filter(!RecordID %in% duplicate_reps_in_sample$RecordID) %>% 
  rowwise() %>% 
  mutate(Genus = str_trim(Genus))  %>%

  mutate(TAXON_EXPERT = case_when(
    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 ~ plant.species, # 4 has a viola split
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 6 ~ plant.species, # has a delphinium split
    sample.id == 6 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',

@@

    sample.id == 18 ~ plant.species,
    sample.id == 19 ~ plant.species,
    sample.id == 20 ~ plant.species,
    sample.id == 21 ~ plant.species,
    sample.id == 23 ~ plant.species, # has both Viola
    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,
    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,
    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,
    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
  #  sample.id == 27 ~ plant.species, # has both Hydrophyllum
    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,
    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 29 ~ plant.species,
    sample.id == 30 ~ plant.species,
    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 32 ~ plant.species,
    sample.id == 32 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 33 ~ plant.species, # has both Viola

    sample.id == 35 ~ plant.species, # has both Viola
    sample.id == 35 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 36 ~ plant.species,

    sample.id == 39 ~ plant.species, # has both hydrophyllum
    sample.id == 39 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 41 ~ plant.species,
    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata',

    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,
    #sample.id == 44 ~ plant.species, # 44 has both delphinium

    sample.id == 46 ~ plant.species,
    sample.id == 49 ~ plant.species,
    sample.id == 49 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.barbeyi',
  #  sample.id == 51 ~ plant.species, # 51 has a viola split
    sample.id == 51 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 52 ~ plant.species,
    sample.id == 54 ~ plant.species,
    sample.id == 55 ~ plant.species,

    sample.id == 56 ~ plant.species,
    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',

@@

   TRUE ~ TAXON_NEW
  )) %>%

  # use this to sum up the total sequences by original read record
  distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
  group_by(sample.id, plant.species) %>% 
  mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%

  # now reduce to the amount of rows of the proper classification
  ungroup() %>% 
  distinct(sample.id, TAXON_EXPERT, .keep_all = T)

  mutate(TAXON_EXPERTS = case_when(
      TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
      'Cynoglossum officinale', 'Cynoglossum pringlei') ~ 'Mertensia',
      TAXON_NEW %in% c('Caltha leptosepala', 'Caltha palustris', 'Trollius europaeus', 'Trollius', 
                       'Thalictrum thalictroides', 'Aquilegia ecalcarata', 
                       'Thalictrum tuberosum') ~ 'Delphinium',
      TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
      TAXON_NEW == 'Scabrethia scabra' ~ 'Wyethia arizonica',
      TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
      TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', #, 
      TAXON_NEW == 'Taraxacum cucullatum' ~ 'Taraxaum oficinale'
    # original was probably right !!! shows the consequence of bad time estimates.#, 
  #   TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  select(TAXON, TAXON_NEW, TAXON_EXPERTS)

>>>>

rm(corb, week_tab, flr_date_ranges, pollen_sample_info, blst)

@@

######## SAMPLE 55 SHOULD BE SOMEWHERE HERE. 
duplicate_reps_in_sample <- blst_expert %>%

  # identify the duplicate samples
  drop_na(plant.species) %>% 
  group_by(RecordID) %>% 
  distinct(plant.species, .keep_all = T) %>% 
  group_by(sample.id, TAXON_NEW) %>% 
  filter(n() >= 2)

dupe_ids <- duplicate_reps_in_sample %>% pull(RecordID) %>% unique()
dupe_samples <- duplicate_reps_in_sample %>% pull(sample.id) %>% unique()

  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Solomon'
duplicate_reps_in_sample <- duplicate_reps_in_sample %>% 
  ungroup() %>% 
  group_by(sample.id, Genus) %>% 
  mutate(Prcnt_seqs_Reclass = 
           ifelse(n() == 2, Prcnt_seqs/2, sum(Prcnt_seqs)/n_distinct(plant.species)/2 ),
         TAXON_EXPERT = plant.species) %>% 
  distinct(sample.id, TAXON_EXPERT, .keep_all = T)

@@

blst_reclass_expert <- blst_expert %>% 
  filter(!RecordID %in% dupe_ids) %>% 
  distinct() %>% 
  rowwise() %>% 
  mutate(Genus = str_trim(Genus))  %>%

  mutate(TAXON_EXPERT = case_when(
    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 4 ~ plant.species, # 4 has a viola split

    sample.id == 6 ~ plant.species, # has a delphinium split

@@

    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',
    sample.id == 18 ~ plant.species,

    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', 
    sample.id == 19 ~ plant.species,

    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
    sample.id == 20 ~ plant.species,
    sample.id == 21 ~ plant.species,

    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 23 ~ plant.species, # has both Viola

    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,

    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,

    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,

    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 27 ~ plant.species, # has both Hydrophyllum

    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,

    sample.id == 29 ~ plant.species,

    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 30 ~ plant.species,

    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 32 ~ plant.species,

    sample.id == 33 ~ plant.species, # has both Viola

    sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  # 7.99
    sample.id == 35 ~ plant.species, # has both Viola
    sample.id == 36 ~ plant.species,

    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 39 ~ plant.species, # has both hydrophyllum
    sample.id == 40 ~ plant.species,

    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
    sample.id == 41 ~ plant.species,

    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,
    sample.id == 44 ~ plant.species, # 44 has both delphinium

    sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 46 ~ plant.species,

    sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
    sample.id == 49 ~ plant.species,

    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 51 ~ plant.species, # 51 has a viola split

    sample.id == 52 ~ plant.species,
    sample.id == 54 ~ plant.species,

    ######### 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? ####
    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 55 ~ plant.species,

    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
    sample.id == 56 ~ plant.species,

@@

   TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  mutate(
    TAXON_EXPERT = str_replace_all(TAXON_EXPERT, '[.]', ' '),
    TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Arenaria globiflora', 'Eremogone congesta'),
    TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
                           'Lupinus sericeus', TAXON_EXPERT),) %>%

  # use this to sum up the total sequences by original read record
  distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
  group_by(sample.id, TAXON_EXPERT) %>% 
  mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%

  # now reduce to the amount of rows of the proper classification
  ungroup() %>% 
  distinct(sample.id, TAXON_EXPERT, .keep_all = T) %>% # what is going on with sample 56 then?

  # add back on the multiple groups
  bind_rows(., duplicate_reps_in_sample)  %>% 
  arrange(sample.id, -Prcnt_seqs_Reclass)

# now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
b <- blst_reclass_expert %>% 
  filter(sample.id %in% dupe_samples) %>% 
  group_by(sample.id, Family) %>% 
  filter(n() >= 3)
